---
title: "Q10_graphs"
author: "Sarah Goldsmith, Caitlin Hicks Pries"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

````{r message = FALSE}
library(plyr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(stringr)
library(sf)
library(mapview)
library(ggcorrplot)
library(PerformanceAnalytics)
library(lme4)
library(nlme)
library(MuMIn)
library(lmerTest)
library(tidyr)
library(emmeans)
library(ggeffects)
#library(maps)
```

#set working directory
```{r}
#setwd("/Users/sarahgoldsmith/Documents/Dartmouth/Q10_review")

setwd("~/Documents/GitHub/Q10_review")

```

#read in file
```{r}
alldata <- read.csv("Q10_alldata_workingfile2.csv")
```

#Unit conversions
```{r}
#make sure columns that are supposed to be numeric are actually numeric
alldata <- alldata %>% 
  mutate_at(vars(MAP, MAT, latitude1, longitude1, activeLayer, topDepth, bottomDepth, SOC, nitrogenMean, cn, bulkDensityMean,  replicates, pH, redox, percentSand, percentSilt, percentClay, microbialBiomass, Q10, minTemp, maxTemp, duration), ~as.numeric(as.character(.))) 

# unit conversions
unique(alldata$carbonUnits)
#convert units to percent
alldata$carbonPercent <- as.numeric(ifelse(alldata$carbonUnits == "percent", alldata$carbonMean, ifelse(
  alldata$carbonUnits == "mg g-1" | alldata$carbonUnits == "g kg-1", alldata$carbonMean /10, ""
)))
# "Aaltonen2019JEM" reports in "kg m-2", but no BD available

unique(alldata$SOCUnits)
alldata$SOCpercent <- as.numeric(ifelse(alldata$SOCUnits == "percent", alldata$SOC, ifelse(alldata$SOCUnits == "SOC (g kg-1)" | alldata$SOCUnits == "g kg -1", alldata$SOC /10, ifelse(alldata$SOCUnits == "g 100 g-1 soil", alldata$SOC, ""))))

#merge SOC and %C
alldata$carbonPercentAll <- ifelse(!is.na(alldata$SOCpercent), alldata$SOCpercent, alldata$carbonPercent)

unique(alldata$nitrogenUnits)
alldata$nitrogenPercent <- as.numeric (ifelse(alldata$nitrogenUnits == "percent", alldata$nitrogenMean, ifelse(
  alldata$nitrogenUnits == "mg/gSoil" | alldata$nitrogenUnits =="g kg-1", alldata$nitrogenMean /10, ifelse( alldata$nitrogenUnits == "g 100g-1 soil", alldata$nitrogenMean, ""
))))

unique(alldata$microbialBiomassUnits)
#convert to mg/kg?
alldata$microbialBiomass2 <- as.numeric (ifelse(alldata$microbialBiomassUnits == "mg kg-1", alldata$microbialBiomass, ifelse(alldata$microbialBiomassUnits == "ug C g soil-1" | alldata$microbialBiomassUnits == "microg/gSoil", alldata$microbialBiomass /1000, ifelse(alldata$microbialBiomassUnits == "mg g-1", alldata$microbialBiomass *1000, ""
))))


#calculate duration in... days??
alldata$durationDays <- as.numeric(ifelse(alldata$durationUnits == "days", alldata$duration, ifelse(alldata$durationUnits == "hr", alldata$duration /24, ifelse(alldata$durationUnits == "year", alldata$duration *365, ifelse(alldata$durationUnits == "weeks", alldata$duration /7, ifelse(alldata$durationUnits == "months", alldata$duration * 30, ""))))))
#used 30 days/month for conversion

#calculate temperature range
alldata$tempRange <- as.numeric(alldata$maxTemp - alldata$minTemp)

#categorize midDepth into categories? 0-10, 10-30, 30-50, 50-end 
alldata$depthCategory <- ifelse(alldata$midDepth <= 10, "0-10", ifelse(alldata$midDepth >10 & alldata$midDepth <= 30, "10-30", ifelse(alldata$midDepth >30 & alldata$midDepth <= 50, "30-50", "50-end")))

alldata$uniqueSite <- as.factor(paste(alldata$citationKey, alldata$site))
alldata$uniqueID <- as.factor(paste(alldata$site, alldata$midDepth))
alldata$citationKey <- as.factor(alldata$citationKey)
```
#Information about the data 
```{r}
#number of unique studies
length(unique(alldata$citationKey))
# number of papers from/not from review papers
print("number of papers")
length(unique(alldata[which(is.na(alldata$reviewPaperKey)),1]))
length(unique(alldata[which(alldata$reviewPaperKey == "Li2020AdvSci"),1]))
length(unique(alldata[which(alldata$reviewPaperKey == "Ren2020GBC"),1]))

# number of data points from/not from review papers
print("number of data points")
length(alldata[which(is.na(alldata$reviewPaperKey)),1])
length(alldata[which(alldata$reviewPaperKey == "Li2020AdvSci"),1])
length(alldata[which(alldata$reviewPaperKey == "Ren2020GBC"),1])


```

#Geography of studies
```{r}
###### map study locations #####
worldmap <- map_data(map = "world")
ggplot() + geom_point(data = alldata, aes(x = longitude1, y = latitude1), color = "black", size = 1) + geom_polygon(data = worldmap, aes(x=long, y = lat, group = group)) + geom_point(data = alldata, aes(x = longitude1, y = latitude1), color = "yellow", size = 0.5)

#number of countries represented 
unique(alldata$country)
length(unique(alldata$country))

ggplot(alldata) + geom_bar(aes(country)) #this might not be the most useful. But most of the data is from China

```

#Exploratory graphs
##Depth
```{r warning=FALSE}

#histogram of Q10 values
ggplot(alldata) + geom_histogram(aes(Q10))

# depth plots
ggplot(alldata) + geom_histogram(aes(midDepth))

ggplot(alldata) + geom_point(aes(y = midDepth, x = Q10))

```

##Environmental conditions
```{r warning=FALSE}
#plot Q10 vs % carbon and nitrogen
ggplot() + geom_point(data = alldata, aes(x = carbonPercentAll, y = Q10))
ggplot() + geom_point(data = alldata, aes(x = SOCpercent, y = Q10))
ggplot() + geom_point(data = alldata, aes(x = nitrogenPercent, y = Q10))
ggplot() + geom_point(data = alldata, aes(x = pH, y = Q10))

ggplot(alldata) + geom_point(aes(x = cn, y = Q10))

#climate variables (temperature, elevation)
ggplot(alldata) + geom_point(aes(x = MAP, y = Q10))
ggplot(alldata) + geom_point(aes(x = MAT, y = Q10))

#sand/silt/clay
ggplot(alldata) + geom_point(aes(x = percentSand, y = Q10))
ggplot(alldata) + geom_point(aes(x = percentSilt, y = Q10))
ggplot(alldata) + geom_point(aes(x = percentClay, y = Q10))

#make facet graphs
ggplot(data = alldata, aes(Q10, midDepth, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~landCover)
#ggsave(filename = "Q10 by landcover.png")

ggplot(data = alldata, aes(Q10, midDepth, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~soilOrder)

ggplot(data = alldata, aes(Q10, carbonPercentAll, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, cn, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, nitrogenPercent, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, microbialBiomass2, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, pH, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, MAT, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)
ggplot(data = alldata, aes(Q10, MAP, color = citationKey)) + geom_point(show.legend = FALSE) + facet_wrap(~depthCategory)

```

##Experimental conditions
```{r warning=FALSE}
#duration plots
ggplot(alldata) + geom_histogram(aes(durationDays))

ggplot(alldata) + geom_point(aes(x = durationDays, y = Q10)) 

#minTemp
ggplot(alldata) + geom_histogram(aes(minTemp))

ggplot(alldata) + geom_point(aes(x = minTemp, y = Q10)) 

# incubation temperature 
ggplot(data = alldata) + geom_point(aes(x = maxTemp, y = Q10))
ggplot(data = alldata) + geom_point(aes(x = minTemp, y = Q10))
ggplot(data = alldata) + geom_point(aes(x = tempRange, y = Q10))

#facet graphs by depth category

ggplot(data = alldata, aes(tempRange, Q10)) + 
  geom_point() + 
  geom_smooth() + 
  facet_wrap(~depthCategory)

ggplot(data = alldata, aes(minTemp, Q10)) + 
  geom_point() + 
  geom_smooth() + 
  facet_wrap(~depthCategory)

ggplot(data = alldata, aes(durationDays, Q10)) + 
  geom_point() + 
  facet_wrap(~depthCategory)

```

##Correlation
```{r warning=FALSE}
#correlation plots
alldata.climate <- alldata %>%
  select(Q10,midDepth, MAT, MAP, pH, cn, carbonPercentAll, nitrogenPercent, microbialBiomass2)
chart.Correlation(alldata.climate, pch = 19)

alldata.incubation <- alldata %>%
  select(Q10,midDepth, minTemp, maxTemp, tempRange, durationDays)
chart.Correlation(alldata.incubation)

```

#Build Models
#Experiment characteristics 
##Model
```{r}

length(which(!is.na(alldata$midDepth) & !is.na(alldata$minTemp)& !is.na(alldata$durationDays)))  #953

#look for optimal random effects
mod<-gls(data=alldata, Q10~ 1+ minTemp + durationDays, na.action = na.omit)
mod2<-lmer(data=alldata, Q10~1+ minTemp + durationDays + (1|uniqueSite), na.action = na.omit)
mod3<-lmer(data=alldata, Q10~1+ minTemp + durationDays + (1|citationKey), na.action = na.omit)
mod4<-lmer(data=alldata, Q10~1+ minTemp + durationDays + (1|citationKey/uniqueSite), na.action = na.omit)
AICc(mod, mod2, mod3, mod4)

summary(mod4)
anova(mod4)

#graphical exploration of residuals
F_Final <- fitted(mod4)
R_Final <- residuals(mod4, type = "pearson", scaled = TRUE) #type="response" 

N = !is.na(alldata$durationDays)
Rfull <- NA
Rfull[N] <- R_Final

op <- par(mfrow = c(2,2), mar = c(5,4,1,1))
plot(F_Final, R_Final) #variance increases with value of Q10
hist(Rfull)
plot(Rfull ~ alldata$minTemp)
plot(Rfull ~ alldata$durationDays)
plot(Rfull ~ alldata$minTemp)
plot(Rfull ~ alldata$maxTemp)
par(op)

#transform response variable
alldata$Q10T<-log(alldata$Q10)

#look for optimal random effects
mod<-gls(data=alldata, Q10T~ 1+ minTemp*midDepth + durationDays*midDepth, na.action = na.omit)
mod2<-lmer(data=alldata, Q10T~1+ minTemp*midDepth + durationDays*midDepth + (1|uniqueSite), na.action = na.omit)
mod3<-lmer(data=alldata, Q10T~1+ minTemp*midDepth + durationDays*midDepth + (1|citationKey), na.action = na.omit)
mod4<-lmer(data=alldata, Q10T~1+ minTemp*midDepth + durationDays*midDepth + (1|citationKey/uniqueSite), na.action = na.omit)
AIC(mod, mod2, mod3, mod4)

summary(mod4)
anova(mod4)

F_Final <- fitted(mod4)
R_Final <- residuals(mod4, type = "pearson", scaled = TRUE) #type="response" 


N = !is.na(alldata$durationDays)
Rfull <- NA
Rfull[N] <- R_Final

op <- par(mfrow = c(2,2), mar = c(5,4,1,1))
plot(F_Final, R_Final) #variance increases with value of Q10
hist(Rfull)
plot(Rfull ~ alldata$minTemp)
plot(Rfull ~ alldata$durationDays)
plot(Rfull ~ alldata$minTemp)
plot(Rfull ~ alldata$tempRange)
par(op)


```

##View the trends of the interaction
```{r}
library(emmeans)

#depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4,~midDepth,at=mylist, CIs=TRUE)

#duration by depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4,~midDepth*durationDays,at=mylist, CIs=TRUE)

dura <- round(mean(alldata$durationDays, na.rm=T) + sd(alldata$durationDays, na.rm=T))
dur <- round(mean(alldata$durationDays, na.rm=T))
durb <- 2

mylist <- list(midDepth=c(5,15,30,50,100),durationDays=c(dura,dur,durb)) 
emmip(mod4,midDepth~durationDays,at=mylist)

#min temperature by depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4,~midDepth*minTemp,at=mylist, CIs=TRUE)

mta <- round(mean(alldata$minTemp, na.rm=T) + sd(alldata$minTemp, na.rm=T))
mtr <- round(mean(alldata$minTemp, na.rm=T))
mtb <- round(mean(alldata$minTemp, na.rm=T) - sd(alldata$minTemp, na.rm=T))

mylist <- list(midDepth=c(5,15,30,50,100),minTemp=c(mta,mtr,mtb)) 
emmip(mod4,midDepth~minTemp,at=mylist)

```

#Environmental Characteristics
##Model
```{r}
length(which(!is.na(alldata$midDepth) & !is.na(alldata$landCover)))  #954
length(which(!is.na(alldata$midDepth) & !is.na(alldata$landCover)& !is.na(alldata$MAP)& !is.na(alldata$MAT))) #835

#look for optimal random effects using transformed Q10
moden <- gls(data=alldata, Q10T ~ 1 + midDepth*landCover + MAP*midDepth + MAT*midDepth , na.action = na.omit)
mod2en <- lmer(data=alldata, Q10T ~ 1 + midDepth*landCover + MAP*midDepth + MAT*midDepth  + (1|uniqueSite), na.action = na.omit)
mod3en <- lmer(data=alldata, Q10T ~ 1 +midDepth*landCover + MAP*midDepth + MAT*midDepth  + (1|citationKey), na.action = na.omit)
mod4en <- lmer(data=alldata,Q10T ~ 1 + midDepth*landCover + MAP*midDepth + MAT*midDepth  + (1|citationKey/uniqueSite), na.action = na.omit)
#Some predictor variables are on very different scales: consider rescaling

AICc(moden, mod2en, mod3en, mod4en)
#lowest AICc = mod4

summary(mod4en)
anova(mod4en)

#graphical exploration of residuals
F_Final <- fitted(mod4en)
R_Final <- residuals(mod4en, type = "pearson", scaled = TRUE) #type="response" 
#no difference in graphs with pearson vs response

N = !is.na(alldata$Q10T)
Rfull <- NA
Rfull[N] <- R_Final

op <- par(mfrow = c(2,2), mar = c(5,4,1,1))
plot(F_Final, R_Final) #variance increases with value of Q10
hist(Rfull)
plot(Rfull ~ alldata$midDepth)
boxplot(Rfull ~ alldata$landCover)
plot(Rfull ~ alldata$MAP)
plot(Rfull ~ alldata$MAT)
plot(Rfull ~ alldata$pH)
plot(Rfull ~ alldata$nitrogenPercent) #decrease in variance with N?
par(op)


```

##View the significant trends 
```{r}

#depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4en,~midDepth,at=mylist, CIs=TRUE)

#LandCover by depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4en,~midDepth|landCover,at=mylist, CIs=TRUE)

emtrends(mod4en, pairwise~ landCover, var = "midDepth")

#MAT partial residual plot

library(ggeffects)
pr<-ggpredict(mod4en, "MAT [all]")
plot(pr, add.data = TRUE)
plot(pr, residuals = TRUE, residuals.line = TRUE) #is absolutely a linear effect

pr<-ggpredict(mod4en, c("midDepth", "landCover"))
plot(pr, add.data = TRUE, grid=TRUE)
plot(pr, residuals = TRUE, residuals.line = TRUE, grid=TRUE) 

```

#Soil characteristics
##Model
Note that I chose C:N here instead of N
```{r}
print("number of data points with all variables (soil model: C, N, pH)")
length(which(!is.na(alldata$midDepth) & !is.na(alldata$cn) & !is.na(alldata$carbonPercentAll)& !is.na(alldata$pH))) #674

#look for optimal random effects using transformed Q10
mod <- gls(data=alldata, Q10T ~ 1 + midDepth*cn + carbonPercentAll*midDepth + pH*midDepth , na.action = na.omit)
mod2 <- lmer(data=alldata, Q10T ~ 1 + midDepth*cn + carbonPercentAll*midDepth + pH*midDepth  + (1|uniqueSite), na.action = na.omit)
mod3 <- lmer(data=alldata, Q10T ~ 1 +midDepth*cn + carbonPercentAll*midDepth + pH*midDepth  + (1|citationKey), na.action = na.omit)
mod4 <- lmer(data=alldata,Q10T ~ 1 + midDepth*cn + carbonPercentAll*midDepth + pH*midDepth  + (1|citationKey/uniqueSite), na.action = na.omit)
#Some predictor variables are on very different scales: consider rescaling

AICc(mod, mod2, mod3, mod4)
#lowest AICc = mod3, 4 is overfit

summary(mod4)
anova(mod4)

#graphical exploration of residuals
F_Final <- fitted(mod4)
R_Final <- residuals(mod4, type = "pearson", scaled = TRUE) #type="response" 
#no difference in graphs with pearson vs response

N = !is.na(alldata$Q10T)
Rfull <- NA
Rfull[N] <- R_Final

op <- par(mfrow = c(2,2), mar = c(5,4,1,1))
plot(F_Final, R_Final) #variance increases with value of Q10
hist(Rfull)
plot(Rfull ~ alldata$midDepth)
plot(Rfull ~ alldata$nitrogenPercent) 
plot(Rfull ~ alldata$carbonPercentAll) #decrease in variance with C
plot(Rfull ~ alldata$cn) #decrease in variance with cn
plot(Rfull ~ alldata$pH)
boxplot(Rfull ~ alldata$landCover)
plot(Rfull ~ alldata$MAP)
plot(Rfull ~ alldata$MAT)
par(op)


```

##View the significant trends 
```{r}

#depth
mylist <- list(midDepth=c(5,15,30,50,100))
emmip(mod4,~midDepth,at=mylist, CIs=TRUE)

#cn by depth
mylist <- list(cn=c(5, 15, 30, 50))
emmip(mod4,~cn,at=mylist, CIs=TRUE)

#pH by depth
pHa <- round(mean(alldata$pH, na.rm=T) + sd(alldata$pH, na.rm=T))
pH <- round(mean(alldata$pH, na.rm=T))
pHb <- round(mean(alldata$pH, na.rm=T) - sd(alldata$pH, na.rm=T))

mylist <- list(midDepth=c(5,15,30,50,100),pH=c(pHa,pH,pHb)) 
emmip(mod3,midDepth~pH,at=mylist)

```
#Figures
  
```{r,eval=FALSE}
#Create own theme for figures
own_theme <- theme_bw(base_size = 11) +
  theme(rect = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        axis.line = element_line(color = "black"),
        panel.grid.minor = element_blank())

#By landCover
pred_Q10_midDepth <- ggpredict(mod4en, terms = c("midDepth [all]", "landCover"), ci.lvl = 0.95)

pred_Q10_midDepth$landCover <- pred_Q10_midDepth$group
pred_Q10_midDepth <- pred_Q10_midDepth %>% 
  mutate(landCover = factor(landCover,
                               levels = c("Forested",
                                          "Rangeland/Grassland",
                                          "Cultivated",
                                          "Tundra",
                                          "Peatland",
                                          "Wetland"),
                               ordered = TRUE))

alldata_4pred <- alldata %>% 
  drop_na(c(Q10T, landCover)) %>% 
  mutate(landCover = factor(landCover,
                               levels = c("Forested",
                                          "Rangeland/Grassland",
                                          "Cultivated",
                                          "Tundra",
                                          "Peatland",
                                          "Wetland"),
                               ordered = TRUE))

#graph with predicted lines on the model showing marginal effect of slope with 
#continuous variables held at their means (0) 

alldata_4pred %>%
ggplot() +
  geom_point(aes(x = midDepth, y = Q10), 
             size = 1.5, alpha = 0.4) +
  geom_ribbon(pred_Q10_midDepth, mapping = aes(x, y=exp(predicted), ymin=exp(conf.low), ymax=exp(conf.high)), alpha=0.3) +
  geom_line(pred_Q10_midDepth, mapping = aes(x, exp(predicted), color = landCover), 
            lwd = 1.5) +
  facet_wrap(~landCover) +
  own_theme +    
  theme(legend.position = "none",
        panel.spacing = unit(.5, "cm", data = NULL)) +
  ylab(expression(paste("Q "[10]))) +
  xlab("Depth (cm)") +
  scale_color_brewer(palette = "BrBG", type="div", direction=6)+
  coord_flip()+
  scale_x_reverse(lim=c(150,0))+
  scale_y_continuous(lim=c(0,11.5))
  

ggsave("Q10_depthXlandCover.jpeg", width=6, height=5.5)

```
